---
title: "Analysis of PR2 metadata"
author: "Lilith South"
date: "2024-10-06"
output: html_document
---

Overall goal of metadata analysis
Make a tree first 
Highlight where the unclassified taxa are on the tree of microsporidia and what their sources are (environment, host, location)
Identify where all microsporidia are by taxonomic information, environment, host

# Annotation: Has the taxonomic information changed
is.true species_old == species 
# Source (Isolate/Environmental/Cultural)
# Environment (Marine/Brackish/Freshwater/Terrestrial/Unknown)
# Host

pr2_sequence_origin <- c(genome, metagenome, transcriptome)
pr2_location_geotype
pr2_location_geoname
pr2_location
pr2_ocean
pr2_sample_type <- c(isolate, environmental, culture, specimen, blanks) # should I fill out the blanks?
  gb_isolation_source
  gb_environmental_sample
  gb_note

gb_host #potentially fill out here if data is lacking? and find how to report host (mammalia, reptile, fish, insecta, crustacea, arachnid, bryozoa, oligochaeta, nematoda, platyhelminth, micro-eukaryotes, mix), from Bojko et al. 2022

1. load required packages
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(rgbif) #to give taxonomic annotations to hosts
library(ggplot2)
library(geodata) #to gain habitat information from gps coordinates from host distribution data
library(terra)
library(geodata)
library(predicts)
library(worrms)
```

2. load in metadata
```{r}
ds <- read.csv2("C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_unchanged.csv", header = TRUE, sep = ",")

View(ds)
```

Create a subset of the dataset with only the columns that I need for metadata analysis
```{r}
sub.ds <- ds %>% 
  select(
    pr2_accession, genbank_accession, species_old, gb_isolation_source, gb_host, gb_environmental_sample, gb_note, pr2_sample_type, pr2_ocean, pr2_location, pr2_sequence_origin, pr2_location_geotype, order, family, species)

View(sub.ds)

str(sub.ds)
```

Some of the columns are repetitive
  Write a loop where for pr2 sample types which are na, if the gb_host is not NA, write "isolate" in   pr2_sample_type

```{r}
sub.ds <- sub.ds %>%
  mutate_all(~ ifelse(. == "", NA, .))

## For the Source column
# put "isolate" if there is an identified host 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$pr2_sample_type[i])) { # check if pr2_sample_type is NA
    if(!is.na(sub.ds$gb_host[i])) { # check if gb_host is not NA
      sub.ds$pr2_sample_type[i] <- "isolate" # assign "isolate" to the specific row
    }
  }
}

# Add a column with environmental type and populate it with "freshwater", "Marine", "Terrestrial", and "unknown"
sub.ds %>% mutate(environment_type = NA)

# put "marine" if there is an identified ocean in pr2_ocean 
sub.ds$environment_type[!is.na(sub.ds$pr2_ocean)] <- "marine" 

# If "nematode" is found in gb_isolation_source, make gb_host "nematode"
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$gb_host[i])) { 
    if(grepl("nematode", sub.ds$gb_isolation_source[i])) { 
      sub.ds$gb_host[i] <- "nematode"
    }
  }
}

# The following to be turned to "freshwater" in environment_type
## any mention of "River", "river", lake", "Lake" "freshwater", "Pond", "pond", "Stream", "swamp" in gb_isolation_source
## "Ammersee", "Pilsensee", "Waldschwaig See", "Woerthsee", "Xiangxi Bay", "Zelivka" in gb_isolation_source
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(!is.na(sub.ds$gb_isolation_source[i]) &&
       (grepl("river|lake|freshwater|pond|stream|swamp|see|River|drinking|cooling|groundwater|tap water|fresh water", 
             sub.ds$gb_isolation_source[i], ignore.case = TRUE) | 
       sub.ds$gb_isolation_source[i] == "Xiangxi Bay" |
       sub.ds$gb_isolation_source[i] == "Zelivka")) { 
      sub.ds$environment_type[i] <- "freshwater"
    }
  }
}

# Other considerations for freshwater
## under pr2_location
### "Loch", "Lagoon", "lake", "Lough Corrib", "river"
### NOT "Mono Lake"
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(!is.na(sub.ds$pr2_location[i]) &&
       (grepl("loch|lake|lagoon|lough|river", 
             sub.ds$pr2_location[i], ignore.case = TRUE) & 
       sub.ds$pr2_location[i] != "Mono Lake")) { 
      sub.ds$environment_type[i] <- "freshwater"
    }
  }
} # Mono Lake is assigned freshwater, I need to make it NA

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(!is.na(sub.ds$pr2_location_geotype[i]) &&
       (grepl("freshwater|lake|river|pond", 
             sub.ds$pr2_location_geotype[i], ignore.case = TRUE))) { 
      sub.ds$environment_type[i] <- "freshwater"
    }
  }
}

for(i in 1:nrow(sub.ds)) {
    if (!is.na(sub.ds$pr2_location[i]) &&
        (sub.ds$pr2_location[i] == "Mono Lake, CA"|
        sub.ds$pr2_location[i] == "Spiral Jetty, Great Salt Lake, UT")) { 
      sub.ds$environment_type[i] <- "salt lake"
    }
  }

# Assign "terrestrial" in environment_type to the following
## "anaerobic full-scale farm reactor treating agricultural wastes", "diapausing larvae in corn stalks", "Moss" in gb_isolation_source
## any mention of "snow", "soil", "bumble bee", "clay", "compost", "field", "silk", "forest", "sand" in gb_isolation_source
# If "Homo sapiens" or "nematode" are the hosts

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(!is.na(sub.ds$gb_isolation_source[i]) &&
       (grepl("sludge|snow|soil|bee|clay|compost|field|silk|forest|sand|Moss|tree",
              sub.ds$gb_isolation_source[i],
              ignore.case = TRUE) | 
       sub.ds$gb_isolation_source[i] == "anaerobic full-scale farm reactor treating agricultural wastes"|
       sub.ds$gb_isolation_source[i] == "diapausing larvae in corn stalks")) {
      sub.ds$environment_type[i] <- "terrestrial"
    }
    
    if (!is.na(sub.ds$gb_host[i]) &&
               (sub.ds$gb_host[i] == "Homo sapiens" |
                sub.ds$gb_host[i] == "nematode")) {
        sub.ds$environment_type[i] <- "terrestrial"
      }
  } 
} 

# Other considerations for terrestrial
## under gb_note
### "ant", "moth", "bee", "rabbit"
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(!is.na(sub.ds$gb_note[i]) &&
       (grepl("ant|moth|bee|rabbitr", 
             sub.ds$gb_note[i], ignore.case = TRUE))) { 
      sub.ds$environment_type[i] <- "terrestrial"
    }
  }
}


# wastewater 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("wastewater|waste water|sewage|sludge|feces", sub.ds$gb_isolation_source[i]))
      { 
      sub.ds$environment_type[i] <- "wastewater"
    }
  }
}

# Other considerations for marine
## under pr2_location
### "Bay", "harbour", "coast", "Puget Sound", "Bay of Fundy, Nova Scotia"
### NOT "South Carolina, Sumter County, Woods Bay State Park", "West Coast of Vancouver Island, British Columbia", "North Bay"

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if((grepl("bay|harbour|coast|Puget Sound", sub.ds$pr2_location[i], ignore.case = TRUE)) &&
       !(sub.ds$pr2_location[i] %in% c(
         "South Carolina, Sumter County, Woods Bay State Park",
         "West Coast of Vancouver Island, British Columbia", 
         "North Bay"
         )))
      { 
      sub.ds$environment_type[i] <- "marine"
    }
  }
}

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("beach|Beach|Harbor|shoreline", sub.ds$gb_isolation_source[i]))
      { 
      sub.ds$environment_type[i] <- "marine"
    }
  }
}

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("beach", sub.ds$pr2_location_geotype[i]))
      { 
      sub.ds$environment_type[i] <- "marine"
    }
  }
}

#For "brackish" in environment_type
## any mention of "brackish", "estuarine" in gb_isolation_source
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("brackish|estuarine", sub.ds$gb_isolation_source[i]))
      { 
      sub.ds$environment_type[i] <- "brackish"
    }
  }
}

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("brackish", sub.ds$pr2_location_geotype[i]))
      { 
      sub.ds$environment_type[i] <- "brackish"
    }
  }
}

# Other considerations for brackish
## under pr2_location
### "estuary", "Eden Landing, San Francisco Bay, CA"
for(i in 1:nrow(sub.ds)) {
    if(grepl("estuary|Eden Landing, San Francisco Bay, CA", sub.ds$pr2_location[i], ignore.case = TRUE))
      { 
      sub.ds$environment_type[i] <- "brackish"
    }
}

# Brackish
for(i in 1:nrow(sub.ds)) {
    if(grepl("estuary|Eden Landing, San Francisco Bay, CA", sub.ds$pr2_location[i], ignore.case = TRUE))
      { 
      sub.ds$environment_type[i] <- "brackish"
    }
}

# "amoebae endoparasite in gb_isolation_source should be "amoebae" in gb_host
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$gb_host[i])) { 
    if(grepl("amoebae", sub.ds$gb_isolation_source[i]))
      { 
      sub.ds$gb_host[i] <- "amoebae"
    }
  }
}


View(sub.ds)

```

# Annotation: Has the taxonomic information changed
is.true species_old == species 
```{r}
sub.ds <- sub.ds %>%
  mutate(annotation_note = if_else(!is.na(species_old) & species_old == species, "improved", "corrected"))

for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$species_old[i])) 
      sub.ds$annotation_note[i] <- "introduced"
    }

# Step 1: Count the occurrences of "the value 5"corrected" in annotation_note
count <- sum(sub.ds[, "annotation_note"] == "corrected", na.rm = TRUE)

# Step 2: Get the total number of rows in the column
total_rows <- nrow(sub.ds)

# Step 3: Calculate the percentage
percentage <- (count / total_rows) * 100

# Print the percentage
print(percentage) #66.90402
```

Annotate the hosts taxonomically
```{r}
# Get taxonomic information for a species
tax_info_test <- name_backbone("Acanthacyclops vernalis
") #one of the hosts to see if it works

# View the result
print(tax_info_test) #example

#create some new columns
sub.ds <- sub.ds %>%
  mutate(host_taxonomy = NA)

for(i in 1:nrow(sub.ds)) {
  if(!(is.na(sub.ds$gb_host[i]))){
    # Get taxonomic information
    host_taxonomy <- tryCatch(
      {name_backbone(name = sub.ds$gb_host[i])
        },
      error = function(e) {
        NULL # Return NULL if there's an error
      }
    )
    
    # Check if the result is not empty and has a 'phylum' field
    if(!is.null(host_taxonomy) && !is.null(host_taxonomy$phylum)) {
      sub.ds$host_taxonomy[i] <- host_taxonomy$phylum
    }
  }
}

#create some new column for host order
sub.ds <- sub.ds %>%
  mutate(host_order = NA)

for(i in 1:nrow(sub.ds)) {
  if(!(is.na(sub.ds$gb_host[i]))){
    # Get taxonomic information
    host_order <- tryCatch(
      {name_backbone(name = sub.ds$gb_host[i])
        },
      error = function(e) {
        NULL # Return NULL if there's an error
      }
    )
    if(!is.null(host_order) && !is.null(host_order$order)) {
      sub.ds$host_order[i] <- host_order$order
    }
  }
}

sub.ds <- sub.ds %>% rename(host_phylum = host_taxonomy)

host_phylum_entries <- unique(sub.ds$host_phylum)
print(host_phylum_entries)
```

```{r}
#Use host species to generate habitat where not specified

for (i in 1:nrow(sub.ds)) {
  
  # Only run query if environment_type is NA
  if (is.na(sub.ds$environment_type[i]) && !is.na(sub.ds$gb_host[i])) {
    
    spname <- sub.ds$gb_host[i]
    message("Querying WoRMS for: ", spname)
    
    res <- tryCatch({
      wm_records_name(name = spname, marine_only = FALSE)
    }, error = function(e) NULL)
    
    if (!is.null(res) && nrow(res) > 0) {
      rec <- res[1, ]  # first match
      
      # Build a collapsed habitat string
      habitats <- c()
      if (!is.na(rec$isMarine) && rec$isMarine == 1) habitats <- c(habitats, "Marine")
      if (!is.na(rec$isBrackish) && rec$isBrackish == 1) habitats <- c(habitats, "Brackish")
      if (!is.na(rec$isFreshwater) && rec$isFreshwater == 1) habitats <- c(habitats, "Freshwater")
      if (!is.na(rec$isTerrestrial) && rec$isTerrestrial == 1) habitats <- c(habitats, "Terrestrial")
      
      # Collapse into one string, e.g. "Marine, Brackish"
      sub.ds$environment_type[i] <- ifelse(length(habitats) > 0, 
                                           paste(habitats, collapse = ", "), 
                                           NA)
    }
  }
}
```

```{r}
#Do the same but with gbif database

for (i in 1:nrow(sub.ds)) {
  
  if (is.na(sub.ds$environment_type[i]) && !is.na(sub.ds$gb_host[i])) {
    
    spname <- sub.ds$gb_host[i]
    message("Querying GBIF for: ", spname)
    
    # Always initialize backbone
    backbone <- NULL
    
    # Try to get backbone record
    backbone <- tryCatch(
      name_backbone(name = spname),
      error = function(e) NULL
    )
    
    # Only proceed if speciesKey exists
    if (!is.null(backbone) && "speciesKey" %in% names(backbone)) {
      key <- backbone$speciesKey
      
      if (!is.na(key)) {
        # Query occurrences
        res <- tryCatch(
          occ_search(taxonKey = key, limit = 20, fields = c("habitat")),
          error = function(e) NULL
        )
        
        if (!is.null(res) && nrow(res$data) > 0) {
          habitats <- unique(na.omit(res$data$habitat))
          
          if (length(habitats) > 0) {
            sub.ds$environment_type[i] <- paste(habitats, collapse = ", ")
          } else {
            sub.ds$environment_type[i] <- NA
          }
        }
      }
    }
  }
}
```


```{r}
# Terrestrial hosts not included 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("Lepidoptera|Primates|Hemiptera|Hymenoptera|Orthoptera
", sub.ds$host_order[i]))
      { 
      sub.ds$environment_type[i] <- "Terrestrial"
    }
  }
}

# Freshwater hosts not included 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("Acanthacyclops vernalis|Cherax|Cottus bairdii|Crangonyx|Daphnia rosea|Diporeia|Lake|Leiarius|Orhtocyclops modustus|Plecoglossus altivelis
", sub.ds$gb_host[i]))
      { 
      sub.ds$environment_type[i] <- "Freshwater"
    }
  }
}

# Terrestrial hosts not included 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("Agapornis|Agrilus anxius|Agriphila|Alabama|Antheraea|Apple|Arion|weevil|Chorthippus|Covasacris|dachshund|Dendroctonus ponderosae|Dichroplus elongatus|Epilachna varivestis
|falcon|fox|Hippodamia|Lamproderma|Liophloeus|Listronotus|maltese|poodle|Mustela nigripes|Oscheius|Oulema melanopus|Phormia regina|Phyllobrotica|Pogona vitticeps|Polypsocus corruptus|pomeranian|Romalea microptera|Solenopsis geminata|Tristira magellanica|Xanthocaecilius sommermanae|Xanthogaleruca luteola
", sub.ds$gb_host[i]))
      { 
      sub.ds$environment_type[i] <- "Terrestrial"
    }
  }
}

# Marine hosts not included 
for(i in 1:nrow(sub.ds)) {
  if(is.na(sub.ds$environment_type[i])) { 
    if(grepl("halibut
|Cypselurus pinnatibarbatus|Pleuronectes vetulus|Lumbriculus|Metapenaeus joineri|Pagrus major|Plumatella sp.|Pygospio elegans|Seriola quinqueradiata|Talitrus sp.


", sub.ds$gb_host[i]))
      { 
      sub.ds$environment_type[i] <- "Marine"
    }
  }
}
```

```{r}
#manually edit the host taxonomy for the entries that missed the taxonomic assignment

write.table(sub.ds,  file = "C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_moved.csv", col.names = TRUE, row.names = FALSE, sep = ",")

```

```{r}
#reload the manually edited .csv file
edit.ds <- read.csv2("C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_moved.csv", header = TRUE, sep = ",")

#create some new column for host class
edit.ds <- edit.ds %>%
  mutate(host_class = NA)

for(i in 1:nrow(edit.ds)) {
  if(!(is.na(edit.ds$gb_host[i]))){
    # Get taxonomic information
    host_class <- tryCatch(
      {name_backbone(name = edit.ds$gb_host[i])
        },
      error = function(e) {
        NULL # Return NULL if there's an error
      }
    )
    
    # Check if the result is not empty and has a 'class' field
    if(!is.null(host_class) && !is.null(host_class$class)) {
      edit.ds$host_class[i] <- host_class$class
    }
  }
}

write.table(edit.ds,  file = "C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_moved.csv", col.names = TRUE, row.names = FALSE, sep = ",")
```


Generate meaningful categories for hosts
```{r}
# did some manual edits to the spreadsheet, reload it

edit.ds <- read.csv2("C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_moved.csv", header = TRUE, sep = ",")

edit.ds <- edit.ds %>% mutate(host_category = NA)

# Amoebozoa <- "Amoebozoa" for host_phylum 
for(i in 1:nrow(edit.ds)) {
  if(is.na(edit.ds$host_category[i])) { # check if host_category is NA
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "Amoebozoa") { 
      edit.ds$host_category[i] <- "Amoebozoa" # assign "Amoebozoa" to the specific row
    }
    {
  if(is.na(edit.ds$host_category[i])) { # check if host_category is NA
    if(!is.na(edit.ds$gb_host[i]) &&
       edit.ds$gb_host[i] == "amoebae") { 
      edit.ds$host_category[i] <- "Amoebozoa" # assign "Amoebozoa" to the specific row
    }
  }
    }
  }
}

# Crustacean <- "Branchiopoda", "Copepoda", "Malacostraca" under host_class
## "Decapoda" under host_order
## "crustacean" under google..subphylum.
for(i in 1:nrow(edit.ds)) {
  if(is.na(edit.ds$host_category[i])) { # check if host_category is NA
    if(edit.ds$host_class[i] %in% c("Branchiopoda", "Copepoda", "Malacostraca")) { 
      edit.ds$host_category[i] <- "Crustacean" 
    }
    
    if(!is.na(edit.ds$host_order[i]) && edit.ds$host_order[i] == "Decapoda"){ 
      edit.ds$host_category[i] <- "Crustacean" 
    }
  }
    if(!is.na(edit.ds$gb_host[i]) && edit.ds$gb_host[i] == "amphipod"){ 
      edit.ds$host_category[i] <- "Crustacean" 
    }
  }

# Fish <- "Anguilliformes", "Atheriniformes", "Aulopiformes", "Beloniformes", "Characiformes", "Clupeiformes", "Cypriniformes", "Gadiformes", "Gasterosteiformes", "Lophiiformes", "Tetraodontiformes", "Scorpaeniformes", "Salmoniformes", "Pleuronectiformes", "Perciformes", "Osmeriformes" in host_order
## "fish" under other.quick.taxonomy.notes
for(i in 1:nrow(edit.ds)) {
  if(is.na(edit.ds$host_category[i])) { # check if host_category is NA
    if(edit.ds$host_order[i] %in% c("Anguilliformes", "Atheriniformes", "Aulopiformes", "Beloniformes", "Characiformes", "Clupeiformes", "Cypriniformes", "Gadiformes", "Gasterosteiformes", "Lophiiformes", "Tetraodontiformes", "Scorpaeniformes", "Salmoniformes", "Pleuronectiformes", "Perciformes", "Osmeriformes")) { 
      edit.ds$host_category[i] <- "Fish" 
    }
  }
}


for(i in 1:nrow(edit.ds)) {
  if(is.na(edit.ds$host_category[i])) { # check if host_category is NA
    # Nematodes <- "Nematoda" under host_phylum
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "Nematoda") { 
      edit.ds$host_category[i] <- "Nematodes"
    }
    # Mammals <- "Mammalia" under host_class
    if(!is.na(edit.ds$host_class[i]) &&
       edit.ds$host_class[i] == "Mammalia") { 
      edit.ds$host_category[i] <- "Mammals"
    }
    # Insects <- "Insecta" under host_class
    if(!is.na(edit.ds$host_class[i]) &&
       edit.ds$host_class[i] == "Insecta") { 
      edit.ds$host_category[i] <- "Insects"
    }
    # Arachnids <- "Arachnida" under host_class
    if(!is.na(edit.ds$host_class[i]) &&
       edit.ds$host_class[i] == "Arachnida") { 
      edit.ds$host_category[i] <- "Arachnids"
    }
   # Birds <- "Aves" under host_class 
    if(!is.na(edit.ds$host_class[i]) &&
       edit.ds$host_class[i] == "Aves") { 
      edit.ds$host_category[i] <- "Birds"
    }
    # Annelids <- "annelida" under host_phylum
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "annelida") { 
      edit.ds$host_category[i] <- "Annelids"
    }
    # Apicomplexa <- "Eugregarinorida" under host_order
    if(!is.na(edit.ds$host_order[i]) && 
       edit.ds$host_order[i] == "Eugregarinorida") { 
      edit.ds$host_category[i] <- "Apicomplexa"
    }
   # Mollusks <- "Mollusca" under host_phylum 
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "Mollusca") { 
      edit.ds$host_category[i] <- "Mollusks"
    }
    # Cilliates <- "Ciliophora" under host_phylum
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "Ciliophora") { 
      edit.ds$host_category[i] <- "Cilliates"
    }
    # Reptiles <- "Testudines", "Squamata" under host_class
    if(!is.na(edit.ds$host_class[i]) &&
       edit.ds$host_class[i] %in% c("Testudines", "Squamata")) { 
      edit.ds$host_category[i] <- "Reptiles" 
    }
    # Bryozoa <- "Bryozoa" under host_phylum
    if(!is.na(edit.ds$host_phylum[i]) &&
       edit.ds$host_phylum[i] == "Bryozoa") { 
      edit.ds$host_category[i] <- "Bryozoa"
    }
    # Cnidaria <- "Cnidaria" under host_phylum
     if(!is.na(edit.ds$host_phylum[i]) &&
        edit.ds$host_phylum[i] == "Cnidaria") { 
      edit.ds$host_category[i] <- "Cnidaria"
     }
    if(!is.na(edit.ds$gb_host[i]) &&
        edit.ds$gb_host[i] == "Actinosporean") { 
      edit.ds$host_category[i] <- "Cnidaria"
    }
  }
}

write.table(edit.ds,  file = "C:/Users/XPS 13 9360/OneDrive - UBC/Sequences/pr2_workshop/finished_propagated_taxonomy_metadata_moved.csv", col.names = TRUE, row.names = FALSE, sep = ",")

#Make manual edits to host_category
```

Yay! now, I need to make the metadata figures and them tie them into the phylogenetic tree (see mic_tree_script.Rmd)


