---
title: "assignTaxonomy_script"
author: "Lilith South"
date: "July 2024"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'root_directory')
```

## Setup:
#### R Packages & Functions

Install packages if needed
```{r}
#install phyloseq
source("https://raw.githubusercontent.com/joey711/phyloseq/master/inst/scripts/installer.R",
       local = TRUE)
install_phyloseq(branch = "devel")
install.packages("tidyverse")
install.packages("phylotools")

# download dada2
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("dada2", version = "3.23")
```

# Load required packages
```{r}
library(phyloseq)
library(tidyverse)
library(phylotools)
library(dada2); packageVersion("dada2")
library(Biostrings); packageVersion("Biostrings")
library(ggplot2)
library(stringr)
library(wacolors)
library(dplyr)
library(seqinr)
library(patchwork)
```

# Sequence dataset from Chauvet et al. 2023

Trim the reference sequences
  filter to 300 minimum length then trim to just V1-V3 using Geneious 
    524 sequences in trimmed alignment
  
```{r}
# Read the reference fasta file
pr2_V1V3_aligned <- "ref_V1V3.fasta"
sequences <- read.fasta(pr2_V1V3_aligned)

# Remove hyphens from the sequences
pr2_V1V3_unaligned <- lapply(sequences, function(x) {
  gsub("-", "", x) 
})

# Write the cleaned sequences to a new file
write.fasta(pr2_V1V3_unaligned, names = names(sequences), file.out = "refV1V3_unaligned.fasta")

# I need to generate a table that has the pr2 accession numbers as one column and the taxonomic information in another separated by ";" using the propagated_taxonomy.xlsx file
# Read and save taxonomy file where I copied and pasted the accession number column and the taxonomy columns 
tax <- read.table(file = "annotation_file.txt", 
                  sep = "\t",
                  header = TRUE)

accession_tax <- unite(tax, "taxonomy", domain:species, sep = ";", remove = TRUE, na.rm = FALSE)

write.table(x=accession_tax,
      file = "11June_accession_tax.txt",
      sep= " ",
      row.names = FALSE,
      quote = FALSE)

# Rename headers of FASTA file
renamed <- rename.fasta(infile = "refV1V3_unaligned.fasta",
             ref_table = accession_tax, 
             outfile = "pr2_train_v1v3_renamed.fasta") 

```

# Dada2 
tutorial https://benjjneb.github.io/dada2/tutorial.html
```{r}
path <- "path" #path is directory to environmental sequences

list.files(path)
```

```{r}
# the forward fastq files are listed as SRRnumber_micropsoridian_diversity_environemnt_metabarcoding_1.fastq and reverse are 2
fnFs <- sort(list.files(path, pattern = "_1.fastq", full.names = TRUE))

fnRs <- sort(list.files(path, pattern = "_2.fastq", full.names = TRUE)) 
# Extract sample names, assuming file names have format: SAMPLENAME_X.fastq
sample.names <- sapply(strsplit(basename(fnFs), "_"), '[', 1)
```

```{r}
plotQualityProfile(fnFs[1:2])
# truncate the forward reads at position 240 (trimming the last 10 nucleotides) 
```

```{r}
#Visualize the reverse reads
plotQualityProfile(fnRs[1:2])
# reverse reads are significantly worse quality, which is common with Illumina sequencing. Truncate at position 175 
```

Filter and trim
```{r}
# Place filtered files in filtered/ subdirectory
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
names(filtFs) <- sample.names
names(filtRs) <- sample.names

out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(240,175),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=FALSE) # On Windows set multithread=FALSE
head(out) # stopped here 9/12/24
```

Learn the Error Rates
```{r}
errF <- learnErrors(filtFs, multithread=TRUE)
# 127302480 total bases in 530427 reads from 2 samples will be used for learning the error rates.
errR <- learnErrors(filtRs, multithread=TRUE)
# 132207075 total bases in 755469 reads from 3 samples will be used for learning the error rates.

# visualize estimated error rates
plotErrors(errF, nominalQ=TRUE)
```

Sample Inference
```{r}
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)

dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
```
dadaFs
  Sample 1 - 261449 reads in 47996 unique sequences.
  Sample 2 - 268978 reads in 97590 unique sequences.
  Sample 3 - 225042 reads in 46975 unique sequences.
  Sample 4 - 330347 reads in 58111 unique sequences.
  Sample 5 - 132280 reads in 44573 unique sequences.
  Sample 6 - 307465 reads in 58158 unique sequences.
  Sample 7 - 241605 reads in 62593 unique sequences.
  Sample 8 - 161286 reads in 50927 unique sequences.
  Sample 9 - 182604 reads in 50021 unique sequences.
  Sample 10 - 156069 reads in 31486 unique sequences.
  Sample 11 - 232056 reads in 67546 unique sequences.
  Sample 12 - 316073 reads in 64004 unique sequences.
  Sample 13 - 296261 reads in 46832 unique sequences.
  Sample 14 - 248159 reads in 40545 unique sequences.
  Sample 15 - 329084 reads in 56281 unique sequences.
  Sample 16 - 262859 reads in 53236 unique sequences.
  Sample 17 - 203020 reads in 44202 unique sequences.


dadaRs
  Sample 1 - 261449 reads in 39906 unique sequences.
  Sample 2 - 268978 reads in 78844 unique sequences.
  Sample 3 - 225042 reads in 38894 unique sequences.
  Sample 4 - 330347 reads in 45358 unique sequences.
  Sample 5 - 132280 reads in 39667 unique sequences.
  Sample 6 - 307465 reads in 53251 unique sequences.
  Sample 7 - 241605 reads in 58828 unique sequences.
  Sample 8 - 161286 reads in 41098 unique sequences.
  Sample 9 - 182604 reads in 47934 unique sequences.
  Sample 10 - 156069 reads in 26547 unique sequences.
  Sample 11 - 232056 reads in 67020 unique sequences.
  Sample 12 - 316073 reads in 57836 unique sequences.
  Sample 13 - 296261 reads in 53302 unique sequences.
  Sample 14 - 248159 reads in 34257 unique sequences.
  Sample 15 - 329084 reads in 48363 unique sequences.
  Sample 16 - 262859 reads in 47590 unique sequences.
  Sample 17 - 203020 reads in 47288 unique sequences.

```{r}
#inspecting the returned dada-class object 
dadaFs[[1]]
```
dada-class: object describing DADA2 denoising results
194 sequence variants were inferred from 47996 input unique sequences.
Key parameters: OMEGA_A = 1e-40, OMEGA_C = 1e-40, BAND_SIZE = 16

```{r}
mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])

# samp1 <- mergers[[1]] 
# samp3 <- mergers[[3]]
```
257730 paired-reads (in 420 unique pairings) successfully merged out of 260661 (in 765 pairings) input.
215032 paired-reads (in 2036 unique pairings) successfully merged out of 267278 (in 5243 pairings) input.
222305 paired-reads (in 359 unique pairings) successfully merged out of 224689 (in 496 pairings) input.
322506 paired-reads (in 477 unique pairings) successfully merged out of 329942 (in 588 pairings) input.
108788 paired-reads (in 789 unique pairings) successfully merged out of 130822 (in 1622 pairings) input.
62062 paired-reads (in 432 unique pairings) successfully merged out of 307086 (in 1120 pairings) input.
221182 paired-reads (in 1067 unique pairings) successfully merged out of 240588 (in 2430 pairings) input.
150393 paired-reads (in 1936 unique pairings) successfully merged out of 159934 (in 2551 pairings) input.
55539 paired-reads (in 193 unique pairings) successfully merged out of 180831 (in 833 pairings) input.
132405 paired-reads (in 364 unique pairings) successfully merged out of 154188 (in 862 pairings) input.
205213 paired-reads (in 2127 unique pairings) successfully merged out of 230942 (in 5317 pairings) input.
88828 paired-reads (in 490 unique pairings) successfully merged out of 315548 (in 886 pairings) input.
259367 paired-reads (in 914 unique pairings) successfully merged out of 295480 (in 1751 pairings) input.
244877 paired-reads (in 539 unique pairings) successfully merged out of 246727 (in 794 pairings) input.
179876 paired-reads (in 889 unique pairings) successfully merged out of 328622 (in 1731 pairings) input.
256264 paired-reads (in 671 unique pairings) successfully merged out of 261943 (in 1067 pairings) input.
20434 paired-reads (in 274 unique pairings) successfully merged out of 202328 (in 1138 pairings) input.

```{r}
# make asv table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
```
[1]    17 12112

```{r}
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
```
sequence length makes sense

Before removing chimeras, I need to remove the primer sequences that are degenerate
```{r}
# remove chimeras
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
```
Identified 9966 bimeras out of 12112 input sequences.
[1]   17 2146

```{r}
sum(seqtab.nochim)/sum(seqtab)
```
[1] 0.9302451
abundance of chimeric sequence varients account for about 7% of the merged sequences 

track reads through pipeline
```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track) # the merged sequences do not decrease too much

View(seqtab.nochim)
saveRDS(seqtab.nochim, "seqtab_nochim.rds")
```

Time to assign taxonomy!
```{r}
seqs <- readRDS("seqtab_nochim.rds")

taxa_assigned_pr2 <- assignTaxonomy(seqs, "pr2_train_v1v3_renamed.fasta",
                       minBoot = 80,
                       multithread = FALSE)
```

```{r}
taxa.print <- taxa_assigned_pr2 # Removing sequence rownames for display only
rownames(taxa.print) <- NULL
#Change the header names to domain, division, subdivision, class, order, family, genus, species 
colnames(taxa_assigned_pr2) = c("domain", "supergroup", "division", "subdivision", "class", "order", "family", "genus", "species")

View(taxa_assigned_pr2)

write.csv(taxa_assigned_pr2, file = "11June_taxa_pr2currated.csv", row.names = TRUE, col.names = TRUE) #I will change the NAs to show higher level taxonomic information in the phyloseq object below

```

Time to make figures- relative abundances
```{r}
#I have issues with seqtab_nochim saving improperly and it is causing my assigned taxa to be wrong, so I'll load this object back into R since I saved it 

seqs <- readRDS("seqtab_nochim.rds")
## I tried reading this csv in, but had issues when making the phyloseq object 

sample_info <- read.table("sample_info.csv",
                          header = TRUE,
                          sep = ",")

rownames(sample_info) <- sample_info[,1]
sample_info <- sample_info[,-1]

ps <- phyloseq(otu_table(seqs, taxa_are_rows = FALSE),
               tax_table(taxa_assigned_pr2),
               sample_data(sample_info))

```

From pipeline "It is more convenient to use short names for our ASVs (e.g. ASV21) rather than the full DNA sequence when working with some of the tables and visualizations from phyloseq, but we want to keep the full DNA sequences for other purposes like merging with other datasets or indexing into reference databases like the Earth Microbiome Project. For that reason we’ll store the DNA sequences of our ASVs in the refseq slot of the phyloseq object, and then rename our taxa to a short string. That way, the short new taxa names will appear in tables and plots, and we can still recover the DNA sequences corresponding to each ASV as needed with refseq(ps)."

```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps
```

```{r}
tax.clean <- ps@tax_table

tax.clean[is.na(tax.clean)] <- ""

for (i in 1:9){ tax.clean[,i] <- as.character(tax.clean[,i])}

####### Fill holes in the tax table while preserving existing '_X'
# separate the if loops

# to add "_X"
for (i in 1:nrow(tax.clean)){
  
  # Fill for supergroup level if missing
  if (tax.clean[i, 2] == "") {
    if (!str_detect(tax.clean[i, 1], "_X$")) {
      tax.clean[i, 2] <- paste(tax.clean[i, 1], "_X", sep = "")
    } 
  } else if (tax.clean[i, 3] == "") {
    # Fill for division level if missing
    if (!str_detect(tax.clean[i, 2], "_X$")) {
      tax.clean[i, 3] <- paste(tax.clean[i, 2], "_X", sep = "")
    } 
  } else if (tax.clean[i, 4] == "") {
    # Fill for subdivision level if missing
    if (!str_detect(tax.clean[i, 3], "_X$")) {
      tax.clean[i, 4] <- paste(tax.clean[i, 3], "_X", sep = "")
    } 
  } else if (tax.clean[i, 5] == "") {
    # Fill for class level if missing
    if (!str_detect(tax.clean[i, 4], "_X$")) {
      tax.clean[i, 5] <- paste(tax.clean[i, 4], "_X", sep = "")
    } 
  } else if (tax.clean[i, 6] == "") {
    # Fill for order level if missing
    if (!str_detect(tax.clean[i, 5], "_X$")) {
      tax.clean[i, 6] <- paste(tax.clean[i, 5], "_X", sep = "")
    } 
  } else if (tax.clean[i, 7] == "") {
    # Fill for family level if missing
    if (!str_detect(tax.clean[i, 6], "_X$")) {
      tax.clean[i, 7] <- paste(tax.clean[i, 6], "_X", sep = "")
    } 
  } else if (tax.clean[i, 8] == "") {
    # Fill for genus level if missing
    if (!str_detect(tax.clean[i, 7], "_X$")) {
      tax.clean[i, 8] <- paste(tax.clean[i, 7], "_X", sep = "")
    } 
  } else if (tax.clean[i, 9] == "") {
    if (!str_detect(tax.clean[i, 8], "_X$")) {
    tax.clean[i, 9] <- paste(tax.clean[i,8], "_sp.", sep = "")
  }
  }
}

# for the rest
for (i in 1:nrow(tax.clean)){
  
  # Fill for supergroup level if missing
  if (tax.clean[i, 3] == "") {
    # supergroup <- tax.clean[i, 2]
    tax.clean[i, 3] <- paste(tax.clean[i, 2], "X", sep = "")
    tax.clean[i, 4] <- paste(tax.clean[i, 2], "XX", sep = "")
    tax.clean[i, 5] <- paste(tax.clean[i, 2], "XXX", sep = "")
    tax.clean[i, 6] <- paste(tax.clean[i, 2], "XXXX", sep = "")
    tax.clean[i, 7] <- paste(tax.clean[i, 2], "XXXXX", sep = "")
    tax.clean[i, 8] <- paste(tax.clean[i, 2], "XXXXXX", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 2], "XXXXXX_sp.", sep = "")
    
    } else if (tax.clean[i, 4] == "") {
    # division <- tax.clean[i, 3]
    tax.clean[i, 4] <- paste(tax.clean[i, 3], "X", sep = "")
    tax.clean[i, 5] <- paste(tax.clean[i, 3], "XX", sep = "")
    tax.clean[i, 6] <- paste(tax.clean[i, 3], "XXX", sep = "")
    tax.clean[i, 7] <- paste(tax.clean[i, 3], "XXXX", sep = "")
    tax.clean[i, 8] <- paste(tax.clean[i, 3], "XXXXX", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 3], "XXXXX_sp.", sep = "")
    
    } else if (tax.clean[i, 5] == "") {
    # subdivision <- tax.clean[i, 4]
    tax.clean[i, 5] <- paste(tax.clean[i, 4], "X", sep = "")
    tax.clean[i, 6] <- paste(tax.clean[i, 4], "XX", sep = "")
    tax.clean[i, 7] <- paste(tax.clean[i, 4], "XXX", sep = "")
    tax.clean[i, 8] <- paste(tax.clean[i, 4], "XXXX", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 4], "XXXX_sp.", sep = "")
    
    } else if (tax.clean[i, 6] == "") {
     # class <- tax.clean[i, 5]
    tax.clean[i, 6] <- paste(tax.clean[i, 5], "X", sep = "")
    tax.clean[i, 7] <- paste(tax.clean[i, 5], "XX", sep = "")
    tax.clean[i, 8] <- paste(tax.clean[i, 5], "XXX", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 5], "XXX_sp.", sep = "")
    
    } else if (tax.clean[i, 7] == "") {
    # order <- tax.clean[i, 6]
    tax.clean[i, 7] <- paste(tax.clean[i, 6], "X", sep = "")
    tax.clean[i, 8] <- paste(tax.clean[i, 6], "XX", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 6], "XX_sp.", sep = "")
     
    } else if (tax.clean[i, 8] == "") {
    # family <- tax.clean[i, 7]
    tax.clean[i, 8] <- paste(tax.clean[i, 7], "X", sep = "")
    tax.clean[i, 9] <- paste(tax.clean[i, 7], "X_sp.", sep = "")
    
    } else if (tax.clean[i, 9] == "") {
      # genus <- tax.clean[i, 8]
    tax.clean[i, 9] <- paste(tax.clean[i,8], "_sp.", sep = "")
  }
} 

# Convert back to tax_table
tax_table(ps) <- as.matrix(tax.clean) 

View(ps@tax_table)
```

Test relative abundance by (1) genus, and also test by (2) old pr2 (and (3) paper's reference database since our sequence cleaning uses a different protocol)

phyloseq-class experiment-level object
otu_table()   OTU Table:         [ 12 taxa and 2 samples ]
sample_data() Sample Data:       [ 2 samples by 2 sample variables ]
tax_table()   Taxonomy Table:    [ 12 taxa by 9 taxonomic ranks ]

```{r}
# To simplify the GP data
ps.pruned = prune_taxa(taxa_sums(ps) > 1000, ps)

total = median(sample_sums(ps.pruned))

standf = function(x, t=total) round(t * (x / sum(x)))

normalized_ps = transform_sample_counts(ps.pruned, standf)

normalized_ps_genus1 = tax_glom(normalized_ps, "genus", NArm = TRUE)
normalized_ps_genus1

normalized_ps_genus1_relabun = transform_sample_counts(normalized_ps_genus1, function(OTU) OTU/sum(OTU) * 100)

normalized_ps_genus1@tax_table

```
order              family                genus                  
"Enterocytozoonida""Mrazekiidae"         "Endoreticulatus" 
"Enterocytozoonida""Mrazekiidae"         "Helmichia"
"Enterocytozoonida""Mrazekiidae"         "Vittaforma"
"Enterocytozoonida""Mrazekiidae"         "Crispospora"
"Enterocytozoonida""Mrazekiidae"         "Mrazekiidae_X"
"Enterocytozoonida""Enterocytozoonida_X" "Anisofilariata"
"Enterocytozoonida""Enterocytozoonida_X" "Globulispora"
"Enterocytozoonida""Enterocytozoonida_X" "Orthosomella"
"Enterocytozoonida""Enterocytozoonida_X" "Enteropsectra"
"Enterocytozoonida""Enterocytozoonida_X" "Enterocytozoonida_XX"
"Amblyosporida"    "Gurleyidae"          "Conglomerata" 
"Amblyosporida"    "Gurleyidae"          "Octosporea"
"Amblyosporida"    "Gurleyidae"          "Pseudoberwaldia"        
"Amblyosporida"    "Gurleyidae"          "Hazardia"            
"Amblyosporida"    "Gurleyidae"          "Gurleyidae_X"          
"Neopereziida"     "Neopereziida_X"      "Neopereziida_XX"                
"Eumicrosporea_X"  "Eumicrosporea_XX"    "Eumicrosporea_XXX"


Grouping by environment type to be easier to read
```{r}
# Merge sample by their sampling environment
# Here we have 2 "samples" (location)
merged_ps = merge_samples(normalized_ps, "environment")
merged_ps

# Group by genus
merged_ps_genus = tax_glom(merged_ps, "genus", NArm = TRUE)
merged_ps_genus

merged_ps_genus_relabun = transform_sample_counts(merged_ps_genus, function(OTU) OTU/sum(OTU) * 100)

genera <- c("Conglomerata", "Hazardia", "Octosporea", "Pseudoberwaldia", "Gurleyidae_X", "Crispospora", "Endoreticulatus", "Helmichia", "Vittaforma", "Mrazekiidae_X", "Anisofilariata", "Globulispora", "Enteropsectra", "Orthosomella",  "Enterocytozoonida_XX", "Neopereziida_XX", "Eumicrosporea_XXX")
```

```{r}
p2_new <- psmelt(merged_ps_genus_relabun) %>%
  ggplot(aes(x = Sample, y = Abundance, fill = factor(genus, genera))) +
  scale_fill_manual(values = c(
   "Conglomerata" = "#062337",
   "Hazardia" = "#0B496F",
    "Octosporea" = "#1170AA", 
    "Pseudoberwaldia" = "#6CABD2",
    "Gurleyidae_X" = "#C7E6F9",
   "Crispospora" = "#173832",
   "Endoreticulatus" = "#29665C", 
    "Helmichia" = "#44AA99", 
    "Vittaforma" = "#7CCBBE",
   "Mrazekiidae_X" = "#D3EEE9",
    "Anisofilariata" = "#F4E76E",
    "Globulispora" = "#60496F",
    "Enteropsectra" = "#2F3F27",
    "Orthosomella" = "#556F44",
    "Enterocytozoonida_XX" = "#F7FFF6",
    "Neopereziida_XX" = "#F7C6B6", 
    "Eumicrosporea_XXX" = "white"
    ), 
   labels = c(
   "Conglomerata" = "Conglomerata",
   "Hazardia" = "Hazardia",
    "Octosporea" = "Octosporea", 
    "Pseudoberwaldia" = "Pseudoberwaldia",
    "Gurleyidae_X" = "Unclassified",
   "Crispospora" = "Crispospora",
   "Endoreticulatus" = "Endoreticulatus", 
    "Helmichia" = "Helmichia", 
    "Vittaforma" = "Vittaforma",
   "Mrazekiidae_X" = "Unclassified",
    "Anisofilariata" = "Anisofilariata",
    "Globulispora" = "Globulispora",
   "Enteropsectra" = "Enteropsectra",
    "Orthosomella" = "Orthosomella",
    "Enterocytozoonida_XX" = "Unclassified",
    "Neopereziida_XX" = "Unclassified",
    "Eumicrosporea_XXX" = "Unclassified"
    )) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 10)) +
  labs(y = "Relative Abundance", title = "Curated PR2", fill = "Genus", x = "Environment")

p2_new
```

```{r}
ggsave("rel_abun_genus_env_currated_pr2_sorted.png", 
       plot = p2_new)

ggsave("rel_abun_genus_env_currated_pr2_sorted.svg", 
       plot = p2_new)
```

Run classification again, but with the old pr2 database
```{r}
setwd(dir='directory')

# Generate the taxonomy file for the previous rendition of the pr2 database
old_pr2_tax <- read.csv(file = "old_pr2_tax.csv", 
                  sep = ",",
                  header = TRUE)

accession_old_pr2_tax <- unite(old_pr2_tax, "taxonomy", domain:species, sep = ";", remove = TRUE, na.rm = FALSE)

write.table(x=accession_old_pr2_tax,
      file = "old_pr2_tax.txt",
      sep= " ",
      row.names = FALSE,
      quote = FALSE)

# Rename headers of FASTA file
## this FASTA file should only contain the sequences from pr2_microsporidai_first.fa (excluding outgroups)
## Run this alignment and trim using Geneious
rename.fasta(infile = "microsporidia_v1v3.fasta", 
             ref_table = accession_old_pr2_tax, 
             outfile = "old_pr2_train_set_mic_v1v3_renamed.fasta") 

old_pr2_ref_renamed <- readDNAStringSet("old_pr2_train_set_mic_v1v3_renamed.fasta")

# remove "old_name_" from taxa that were outgroups
old_pr2_ref_sequence_headers <- names(old_pr2_ref_renamed)

filtered_old_pr2_ref_fasta <- old_pr2_ref_renamed[!old_pr2_ref_sequence_headers %in% "old_name_"]

writeXStringSet(filtered_old_pr2_ref_fasta, "filtered_old_pr2_ref.fasta")

View(filtered_old_pr2_ref_fasta)
```

Time to assign taxonomy with the old pr2 database
```{r}
setwd(dir='directory')

seqs <- readRDS("seqtab_nochim.rds")

taxa_assigned_old_pr2 <- assignTaxonomy(seqs, "filtered_old_pr2_ref.fasta",
                       minBoot = 80,
                       multithread = FALSE)

old_taxa.print <- taxa_assigned_old_pr2 # Removing sequence rownames for display only
rownames(old_taxa.print) <- NULL
#Change the header names to domain, division, subdivision, class, order, family, genus, species 
colnames(taxa_assigned_old_pr2) = c("domain", "supergroup", "division", "subdivision", "class", "order", "family", "genus", "species")

write.csv(taxa_assigned_old_pr2, file = "taxa_old_pr2.csv", row.names = TRUE, col.names = TRUE) 
```

```{r}
setwd(dir='directory')
taxa_assigned_old_pr2 <- read.csv("taxa_old_pr2.csv", row.names = 1)

tax_old_pr2_matrix <- as.matrix(taxa_assigned_old_pr2)
tax_table_old_pr2_obj <- tax_table(tax_old_pr2_matrix)

ps_old_pr2 <- phyloseq(otu_table(seqs, taxa_are_rows = FALSE),
               tax_table(tax_table_old_pr2_obj),
               sample_data(sample_info))
```

```{r}
dna_old_pr2 <- Biostrings::DNAStringSet(taxa_names(ps_old_pr2))
names(dna_old_pr2) <- taxa_names(ps_old_pr2)
ps_old_pr2 <- merge_phyloseq(ps_old_pr2, dna_old_pr2)
taxa_names(ps_old_pr2) <- paste0("ASV", seq(ntaxa(ps_old_pr2)))
ps_old_pr2
```

```{r}
tax.clean_old_pr2 <- ps_old_pr2@tax_table

tax.clean_old_pr2[is.na(tax.clean_old_pr2)] <- ""

for (i in 1:9){ tax.clean_old_pr2[,i] <- as.character(tax.clean_old_pr2[,i])}

####### Fill holes in the tax table while preserving existing '_X'
# separate the if loops

# for [i, i] == "_X$"
for (i in 1:nrow(tax.clean_old_pr2)){
  
  # Fill for supergroup level if missing
  if (tax.clean_old_pr2[i, 2] == "") {
    if (!str_detect(tax.clean_old_pr2[i, 1], "_X$")) {
      tax.clean_old_pr2[i, 2] <- paste(tax.clean_old_pr2[i, 1], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 3] == "") {
    # Fill for division level if missing
    if (!str_detect(tax.clean_old_pr2[i, 2], "_X$")) {
      tax.clean_old_pr2[i, 3] <- paste(tax.clean_old_pr2[i, 2], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 4] == "") {
    # Fill for subdivision level if missing
    if (!str_detect(tax.clean_old_pr2[i, 3], "_X$")) {
      tax.clean_old_pr2[i, 4] <- paste(tax.clean_old_pr2[i, 3], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 5] == "") {
    # Fill for class level if missing
    if (!str_detect(tax.clean_old_pr2[i, 4], "_X$")) {
      tax.clean_old_pr2[i, 5] <- paste(tax.clean_old_pr2[i, 4], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 6] == "") {
    # Fill for order level if missing
    if (!str_detect(tax.clean_old_pr2[i, 5], "_X$")) {
      tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 5], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 7] == "") {
    # Fill for family level if missing
    if (!str_detect(tax.clean_old_pr2[i, 6], "_X$")) {
      tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 6], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 8] == "") {
    # Fill for genus level if missing
    if (!str_detect(tax.clean_old_pr2[i, 7], "_X$")) {
      tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 7], "_X", sep = "")
    } 
  } else if (tax.clean_old_pr2[i, 9] == "") {
    if (!str_detect(tax.clean_old_pr2[i, 8], "_X$")) {
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i,8], "_sp.", sep = "")
  }
  }
}

# for the rest
for (i in 1:nrow(tax.clean_old_pr2)){
  
  # Fill for supergroup level if missing
  if (tax.clean_old_pr2[i, 2] == "") {
    #domain <- tax.clean_old_pr2[i, 1]
    tax.clean_old_pr2[i, 2] <- paste(tax.clean_old_pr2[i, 1], "X", sep = "")
    tax.clean_old_pr2[i, 3] <- paste(tax.clean_old_pr2[i, 1], "XX", sep = "")
    tax.clean_old_pr2[i, 4] <- paste(tax.clean_old_pr2[i, 1], "XXX", sep = "")
    tax.clean_old_pr2[i, 5] <- paste(tax.clean_old_pr2[i, 1], "XXXX", sep = "")
    tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 1], "XXXXX", sep = "")
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 1], "XXXXXX", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 1], "XXXXXXX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 1], "XXXXXXX_sp.", sep = "")
    
    } else if (tax.clean_old_pr2[i, 3] == "") {
    # supergroup <- tax.clean_old_pr2[i, 2]
     tax.clean_old_pr2[i, 3] <- paste(tax.clean_old_pr2[i, 2], "X", sep = "")
    tax.clean_old_pr2[i, 4] <- paste(tax.clean_old_pr2[i, 2], "XX", sep = "")
    tax.clean_old_pr2[i, 5] <- paste(tax.clean_old_pr2[i, 2], "XXX", sep = "")
    tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 2], "XXXX", sep = "")
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 2], "XXXXX", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 2], "XXXXXX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 2], "XXXXXX_sp.", sep = "")
    
    } else if (tax.clean_old_pr2[i, 4] == "") {
    # division <- tax.clean_old_pr2[i, 3]
    tax.clean_old_pr2[i, 4] <- paste(tax.clean_old_pr2[i, 3], "X", sep = "")
    tax.clean_old_pr2[i, 5] <- paste(tax.clean_old_pr2[i, 3], "XX", sep = "")
    tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 3], "XXX", sep = "")
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 3], "XXXX", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 3], "XXXXX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 3], "XXXXX_sp.", sep = "")
    
    } else if (tax.clean_old_pr2[i, 5] == "") {
    # subdivision <- tax.clean_old_pr2[i, 4]
    tax.clean_old_pr2[i, 5] <- paste(tax.clean_old_pr2[i, 4], "X", sep = "")
    tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 4], "XX", sep = "")
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 4], "XXX", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 4], "XXXX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 4], "XXXX_sp.", sep = "")
    
    } else if (tax.clean_old_pr2[i, 6] == "") {
     # class <- tax.clean_old_pr2[i, 5]
    tax.clean_old_pr2[i, 6] <- paste(tax.clean_old_pr2[i, 5], "X", sep = "")
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 5], "XX", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 5], "XXX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 5], "XXX_sp.", sep = "")
    
    } else if (tax.clean_old_pr2[i, 7] == "") {
    # order <- tax.clean_old_pr2[i, 6]
    tax.clean_old_pr2[i, 7] <- paste(tax.clean_old_pr2[i, 6], "X", sep = "")
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 6], "XX", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 6], "XX_sp.", sep = "")
     
    } else if (tax.clean_old_pr2[i, 8] == "") {
    # family <- tax.clean_old_pr2[i, 7]
    tax.clean_old_pr2[i, 8] <- paste(tax.clean_old_pr2[i, 7], "X", sep = "")
    tax.clean_old_pr2[i, 9] <- paste(tax.clean_old_pr2[i, 7], "X_sp.", sep = "")
    }
} 

# Convert back to tax_table
tax_table(ps_old_pr2) <- as.matrix(tax.clean_old_pr2) 

View(ps_old_pr2@tax_table)
```


#Grouping by environment type to be easier to read
```{r}
# Merge sample by their sampling environment
# Here we have 2 "samples" (location)
merged_ps_old = merge_samples(normalized_ps_old_pr2, "environment")
merged_ps_old

# Group by genus
merged_ps_old_genus = tax_glom(merged_ps_old, "genus", NArm = TRUE)
merged_ps_old_genus

merged_ps_old_genus_relabun = transform_sample_counts(merged_ps_old_genus, function(OTU) OTU/sum(OTU) * 100)

genera_old_pr2 <- c("Conglomerata", "Hazardia", "Octosporea", "Anostracospora", "Cystosporogenes", "Endoreticulatus", "Helmichia", "Vittaforma", "Anisofilariata", "Globulispora", "Enteropsectra", "Liebermannia", "Orthosomella", "Microsporida_XX", "Fungi_XXXX")
```

```{r}
p1_old <- psmelt(merged_ps_old_genus_relabun) %>%
  ggplot(aes(x = Sample, y = Abundance, fill = factor(genus, genera_old_pr2))) +
  scale_fill_manual(values = c(
     "Conglomerata" = "#062337",
     "Hazardia" = "#0B496F",
    "Octosporea" = "#1170AA",
    "Anostracospora" = "#102722",
    "Cystosporogenes" = "#204F47",
    "Endoreticulatus" = "#29665C", 
    "Helmichia" = "#44AA99", 
    "Vittaforma" = "#7CCBBE", 
    "Anisofilariata" = "#F4E76E",
    "Globulispora" = "#60496F",
    "Enteropsectra" = "#2F3F27",
    "Liebermannia" = "#556F44",
    "Orthosomella" = "#89AA74",
    "Microsporida_XX" = "white",
    "Fungi_XXXX" = "grey"),
    labels = c(
     "Conglomerata" = "Conglomerata",
     "Hazardia" = "Hazardia",
    "Octosporea" = "Octosporea",
    "Anostracospora" = "Anostracospora",
    "Cystosporogenes" = "Cystosporogenes",
    "Endoreticulatus" = "Endoreticulatus", 
    "Helmichia" = "Helmichia", 
    "Vittaforma" = "Vittaforma", 
    "Anisofilariata" = "Anisofilariata",
    "Globulispora" = "Globulispora",
    "Enteropsectra" = "Enteropsectra",
    "Liebermannia" = "Liebermannia",
    "Orthosomella" = "Orthosomella",
    "Microsporida_XX" = "Unclassified",
    "Fungi_XXXX" = "Unclassified")) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 10)) +
  labs(y = "Relative Abundance", title = "Previous PR2", fill = "Genus", x = "Environment") 
# + scale_fill_wa_d(wacolors$rainier)

p1_old
```


```{r}
ggsave("rel_abun_genus_env_old_pr2_sorted.png", 
       plot = rel_abun_by_genus_and_env_old,
       path = "plots")
ggsave("rel_abun_genus_env_old_pr2_sorted.svg", 
       plot = rel_abun_by_genus_and_env_old,
       path = "plots")

```

phyloseq-class experiment-level object
otu_table()   OTU Table:         [ 16 taxa and 17 samples ]
sample_data() Sample Data:       [ 17 samples by 2 sample variables ]
tax_table()   Taxonomy Table:    [ 16 taxa by 9 taxonomic ranks ]
refseq()      DNAStringSet:      [ 16 reference sequences ]

Time to test the Dubuffet database
  I have separate a fasta file alignment and a list of accession numbers with taxonomic information   I need to set the alignment up like I did with my alignment 
```{r}
setwd(dir="directory")

# Read and save taxonomy file
ch_tax <- read.table(file = "Taxonomy.txt", 
                  header = FALSE, 
                  sep = "\t")
View(ch_tax)

```

```{r}
setwd(dir="directory")
#read and save fasta file with trimmed reference sequences as done with pr2 reference sequences

# Rename headers of FASTA file
ch_renamed <- rename.fasta(infile = "microsporidia_eukaryota_v1v3Unpaired.fasta", 
             ref_table = ch_tax, 
             outfile = "chauvet_ref_seqs_v1v3_renamed.fasta") 
```

Time to assign taxonomy according to the Dubuffet database
```{r}
seqs <- readRDS("seqtab_nochim.rds")

taxa_assigned_ch <- assignTaxonomy(seqs, "chauvet_ref_seqs_v1v3_renamed.fasta",
                       minBoot = 80,
                       multithread = FALSE)

# remove columns containing all NAs
taxa_assigned_ch <- taxa_assigned_ch[, apply(taxa_assigned_ch, 2, function(x) !all(is.na(x)))]

#rename "Family" column to be "Assigned Group"
colnames(taxa_assigned_ch)[colnames(taxa_assigned_ch) == "Family"] <- "Assigned Group"

write.csv(taxa_assigned_ch, file = "taxa_ch.csv", row.names = TRUE, col.names = TRUE) 
```

```{r}
## If there are issues loading taxa_assigned, use the path below
# taxa_assigned_ch <- read.csv("taxa_ch.csv", row.names = 1)

tax_ch_matrix <- as.matrix(taxa_assigned_ch)
tax_table_ch_obj <- tax_table(tax_ch_matrix)

ps_ch <- phyloseq(otu_table(seqs, taxa_are_rows = FALSE),
               tax_table(tax_table_ch_obj),
               sample_data(sample_info))
```

```{r}
dna_ch <- Biostrings::DNAStringSet(taxa_names(ps_ch))
names(dna_ch) <- taxa_names(ps_ch)
ps_ch <- merge_phyloseq(ps_ch, dna_ch)
taxa_names(ps_ch) <- paste0("ASV", seq(ntaxa(ps_ch)))
ps_ch
```

```{r}
tax.clean_ch <- ps_ch@tax_table

tax.clean_ch[is.na(tax.clean_ch)] <- ""

for (i in 1:7){ tax.clean_ch[,i] <- as.character(tax.clean_ch[,i])}

####### Fill holes in the tax table while preserving existing '_X'
# separate the if loops

# for [i, i] == "_X$"
for (i in 1:nrow(tax.clean_ch)){
  
  # Fill for supergroup level if missing
  if (tax.clean_ch[i, 2] == "") {
    if (!str_detect(tax.clean_ch[i, 1], "_X$")) {
      tax.clean_ch[i, 2] <- paste(tax.clean_ch[i, 1], "_X", sep = "")
    } 
  } else if (tax.clean_ch[i, 3] == "") {
    # Fill for division level if missing
    if (!str_detect(tax.clean_ch[i, 2], "_X$")) {
      tax.clean_ch[i, 3] <- paste(tax.clean_ch[i, 2], "_X", sep = "")
    } 
  } else if (tax.clean_ch[i, 4] == "") {
    # Fill for subdivision level if missing
    if (!str_detect(tax.clean_ch[i, 3], "_X$")) {
      tax.clean_ch[i, 4] <- paste(tax.clean_ch[i, 3], "_X", sep = "")
    } 
  } else if (tax.clean_ch[i, 5] == "") {
    # Fill for class level if missing
    if (!str_detect(tax.clean_ch[i, 4], "_X$")) {
      tax.clean_ch[i, 5] <- paste(tax.clean_ch[i, 4], "_X", sep = "")
    } 
  } else if (tax.clean_ch[i, 6] == "") {
    # Fill for order level if missing
    if (!str_detect(tax.clean_ch[i, 5], "_X$")) {
      tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 5], "_X", sep = "")
    } 
  } else if (tax.clean_ch[i, 7] == "") {
    if (!str_detect(tax.clean_ch[i, 6], "_X$")) {
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i,6], "_sp.", sep = "")
  }
  }
}

# for the rest
for (i in 1:nrow(tax.clean_ch)){
  
  # Fill for Kingdom level if missing
  if (tax.clean_ch[i, 2] == "") {
    #Kingdom <- tax.clean_ch[i, 1]
    tax.clean_ch[i, 2] <- paste(tax.clean_ch[i, 1], "X", sep = "")
    tax.clean_ch[i, 3] <- paste(tax.clean_ch[i, 1], "XX", sep = "")
    tax.clean_ch[i, 4] <- paste(tax.clean_ch[i, 1], "XXX", sep = "")
    tax.clean_ch[i, 5] <- paste(tax.clean_ch[i, 1], "XXXX", sep = "")
    tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 1], "XXXXX", sep = "")
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i, 1], "XXXXX_sp.", sep = "")
    
    } else if (tax.clean_ch[i, 3] == "") {
    # Phylum <- tax.clean_ch[i, 2]
     tax.clean_ch[i, 3] <- paste(tax.clean_ch[i, 2], "X", sep = "")
    tax.clean_ch[i, 4] <- paste(tax.clean_ch[i, 2], "XX", sep = "")
    tax.clean_ch[i, 5] <- paste(tax.clean_ch[i, 2], "XXX", sep = "")
    tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 2], "XXXX", sep = "")
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i, 2], "XXXX_sp.", sep = "")
    
    } else if (tax.clean_ch[i, 4] == "") {
    # Class <- tax.clean_ch[i, 3]
    tax.clean_ch[i, 4] <- paste(tax.clean_ch[i, 3], "X", sep = "")
    tax.clean_ch[i, 5] <- paste(tax.clean_ch[i, 3], "XX", sep = "")
    tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 3], "XXX", sep = "")
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i, 3], "XXX_sp.", sep = "")
    
    } else if (tax.clean_ch[i, 5] == "") {
    # Order <- tax.clean_ch[i, 4]
    tax.clean_ch[i, 5] <- paste(tax.clean_ch[i, 4], "X", sep = "")
    tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 4], "XX", sep = "")
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i, 4], "XX_sp.", sep = "")
    
    } else if (tax.clean_ch[i, 6] == "") {
     # Assigned Group <- tax.clean_ch[i, 5]
    tax.clean_ch[i, 6] <- paste(tax.clean_ch[i, 5], "X", sep = "")
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i, 5], "X_sp.", sep = "")
    
    } else if (tax.clean_ch[i, 7] == "") {
    tax.clean_ch[i, 7] <- paste(tax.clean_ch[i,6], "_sp.", sep = "")
  }
    } 

# Convert back to tax_table
tax_table(ps_ch) <- as.matrix(tax.clean_ch) 

View(ps_ch@tax_table)
```

All samples grouped by environment
```{r}
# Merge sample by their sampling environment
merged_ps_ch = merge_samples(normalized_ps_ch, "environment")
merged_ps_ch

# Group by genus
merged_ps_ch_genus = tax_glom(merged_ps_ch, "Genus", NArm = TRUE)
merged_ps_ch_genus

merged_ps_ch_genus_relabun = transform_sample_counts(merged_ps_ch_genus, function(OTU) OTU/sum(OTU) * 100)

genera_ch <- c("Conglomerata", "Larssonia", "Octosporea", "Pseudoberwaldia", " Aquatic outgroup_X", "Crispospora", "Endoreticulatus", "Helmichia", "Vittaforma", " Arlequin group_X", "Globulispora", "Enteropsectra", "Liebermannia", "Orthosomella", "Orthosomella-like group_X", "IV2_XX", "Paranosema-like group_X", "Microsporidia_XXX", "Eukaryota_XXXXX")
```

```{r}
p3_ch <- psmelt(merged_ps_ch_genus_relabun) %>%
  ggplot(aes(x = Sample, y = Abundance, fill = factor(Genus, genera_ch))) +
  scale_fill_manual(values = c(
    "Conglomerata" = "#062337", 
    "Larssonia" = "#0C4A71",
    "Octosporea" = "#1170AA", 
    "Pseudoberwaldia" = "#6CABD2",
    " Aquatic outgroup_X" = "#C7E6F9",  
    "Paranosema-like group_X" = "#EE8866",
    "Crispospora" = "#173832",
    "Endoreticulatus" = "#29665C", 
    "Helmichia" = "#44AA99", 
    "Vittaforma" = "#7CCBBE", 
    " Arlequin group_X" = "#D3EEE9", 
    "Globulispora" = "#60496F",
    "Enteropsectra" = "#2F3F27",
    "Liebermannia" = "#556F44",
    "Orthosomella" = "#89AA74",
    "Orthosomella-like group_X" = "#B4C9A6",
    "IV2_XX" = "#F7FFF6",
    "Microsporidia_XXX" = "white"),
    labels = c(
    "Conglomerata" = "Conglomerata", 
    "Larssonia" = "Larssonia",
    "Octosporea" = "Octosporea", 
    "Pseudoberwaldia" = "Pseudoberwaldia",
    " Aquatic outgroup_X" = "Unclassified",  
    "Paranosema-like group_X" = "Unclassified", 
    "Crispospora" = "Crispospora",
    "Endoreticulatus" = "Endoreticulatus", 
    "Helmichia" = "Helmichia", 
    "Vittaforma" = "Vittaforma", 
    " Arlequin group_X" = "Unclassified", 
    "Globulispora" = "Globulispora",
    "Enteropsectra" = "Enteropsectra",
    "Liebermannia" = "Liebermannia",
    "Orthosomella" = "Orthosomella",
    "Orthosomella-like group_X" = "Unclassified",
    "IV2_XX" = "Unclassified",
    "Microsporidia_XXX" = "Unclassified")) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 10)) +
  labs(y = "Relative Abundance", title = "Dubuffet et al. 2021", fill = "Genus", x = "Environment")
# + scale_fill_wa_d(wacolors$rainier)

p3_ch
# Look into changing color for Pancytospora
```

```{r}
ggsave("rel_abun_genus_env_currated_ch_sorted.png", 
       plot = rel_abun_by_genus_and_env_ch,
       path = "plots")
ggsave("rel_abun_genus_env_currated_ch_sorted.svg", 
       plot = rel_abun_by_genus_and_env_ch,
       path = "plots")
```

There are some differences between the pr2 database and the chauvet database, especially with some taxa showing that they are uncharacterized using the pr2 database. I want to see if this is a short coming of my database or the chauvet database. So, I will see what similar reference sequences there are and compare their taxonomic annotations between the two 
## This step had been done and we repeated lines 52-531 

Using Bash, I found that all shared sequences (432) are described species
Now I need to look at what sequences are in Chauvet, but not in pr2...

Read in the taxonomy files and see what taxa aren't included in the pr2
```{r}
ch_mic <- read.table("microsporidia_taxonomy.txt", sep = "\t", header = FALSE)

pr2_mic <- read.table("accession_tax.txt", header = TRUE, stringsAsFactors = FALSE)

# Extract prefixes from pr2_mic
pr2_mic_acc <- pr2_mic %>%
  mutate(pr2_accession_prefix = sub("\\..*", "", pr2_accession))

#Find lines in ch_mic not in pr2_mic
ch_unique <- ch_mic %>%
  anti_join(pr2_mic_acc, by = c("V1" = "pr2_accession_prefix"))

View(ch_unique)
#98 sequences found in the ch database, but not the pr2 database. I will add these sequences to the alignment, assign their taxonomy and redo the taxonomic classification 

```

```{r}
ch_unique_acc <- ch_unique[, 1]

write.table(ch_unique_acc, "unique_taxa.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

Or I could use this list to pull out the sequences from the .fasta file from chauvet and then add those already trimmed sequences to my alignment 
When I do this, there are only 16 taxa from the Chauvet alignment that are not present in the pr2 database

>AKIJ01000013 Nematocida ausubeli
>LTDK01000184 Nematocida ironsii
>KX360147 Nematocida minor
>AEOO01000054 Nematocida parisii
>MN935433 Ameson herrnkindi
>KM657354 Dictyocoela berillonum
>MK568064 Glugea eda
>MT006314 Cambaraspora  ﬂoridanus
>MN969054 Myosporidium ladogensis
>FJ865223_B Mockfordia xanthocaeciliae
>MG725375 Vairimorpha alticae
>MN841279 Vairimorpha leptinotarsae
>MF882996 Vairimorpha neumanni
>KY783624 Nosema maddoxi
>MNPJ01000027 Enterocytozoon hepatopenaei
>LWDP01000010 Enterospora canceri

Include these sequences in another run of trimming to SSU, aligning, tree making, and curating, then taxonomic classification

# Use patchwork to patch the three graphs together
```{r}
classify.fig.patch <- p1_old + p2_new + p3_ch

classify.fig.patch

ggsave("classified_all_datasets.svg", 
       plot = classify.fig.patch,
       width = 7, 
       path = "plots")
```
Save, but continue to edit the .svg file
