---
title: "PR2 Tree"
author: "Lilith South"
date: "2024-10-09"
output: html_document
---

# This R Markdown includes code chunks for visualizing Microsporidia PR2 trees
See link from Corey Holt for some guidance https://github.com/coreyholt/ggbeginners/blob/main/ggbeginners_phylogeneticTrees.R
```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'root_directory') 
```

1. load required libraries     
```{r}
library(tidyverse)
library(ape) # for tree manipulation
library(dplyr) # for data manipulation
library(ggtree)
library(tidytree)
library(ggnewscale) # allow for multiple scale_fills
library(gginnards) # to move layers
library(treeio)
library(writexl)
library(readr)
library(ggplot2)
library(svglite) # used with ggsave for better exporting of svg files
library(ggrepel) # prevents overlapping of text
library(rainbow)
library(phyloseq)
library(phylotools)
library(stringr)
library(ggtext) 
library(wacolors) # colorblind friendly pallet 
library(readxl) # for importing excel files
```

2. Read in tree file
```{r}
#rooted tree
raxml <- read.raxml("RAxML_bipartitionsBranchLabels.tre")

raxml
```
3. vizualize raxml with ggtree
```{r}
tree1 <- ggtree(raxml) 

 tree1
```

4. Rename tips with taxonomic information 
```{r}
# First I need to extract the accession numbers that are actually in the tree from the table of all pr2 accession numbers (and outgroups)

og_lb <- get.tree(raxml)$tip.label

all_pr2_labels <- read.table("annotation 2026June02.txt", header = TRUE, sep = "\t")

og_lb_with_tax <- subset(all_pr2_labels, og_tip_lab %in% og_lb)

View(og_lb_with_tax)
```

# Provide descriptive labels for when taxa do not have species names
I already ran and edited this in the file listed below, so I will open and provide new annotations where needed and re-save in the working directory 
```{r}
# I will place the original gb_definition in another column, and change the new label to the definition for taxa that end in "sp.". I'll then need to curate the names

og_lb_with_tax <- og_lb_with_tax %>%
  rename(pr2_accession = og_tip_lab)

ogMetaTable <- read_excel("finished_propagated_taxonomy_metadata_unchanged.xlsx")

merged_lb <- merge(og_lb_with_tax, ogMetaTable[, c("pr2_accession", "gb_definition")], 
                   by = "pr2_accession", 
                   all.x = TRUE)

write.csv(merged_lb, "merged_lb.csv", row.names = FALSE)

```

# New labels
Be sure to change the file name before uploading back, or you could accidentally rewrite!
```{r}
edit_lb <- read.csv("edited_merged_lb.csv")
```

Determine node positions for taxonomic annotation
```{r}
# Add taxonomic names and identify nodes

## Family
fam_lb_tree <- tree1 %<+% og_lb_with_tax +
  geom_tiplab(linewidth = .5, aes(label = family)) +
  geom_text(size = .5, aes(label=node), color = "forestgreen")+
  geom_tree(size = .25)

## Order
ord_lb_tree <- tree1 %<+% og_lb_with_tax +
  geom_tiplab(linewidth = .5, aes(label = order)) +
  geom_text(size = .5, aes(label=node), color = "forestgreen")+
  geom_tree(size = .25)

ggsave(
  file = "fam_lbl_tree.svg",
  plot = fam_lb_tree)

ggsave(
  file = "order_lbl_tree.svg",
  plot = ord_lb_tree)
```
## Taxonomic Annotation
# Order
"Enterocytozoonida" 750
"Nosema" 852 #To be edited
"Glugeida" 619
"Amblyosporida" 548
"Neopereziida" 919
"Caudosporida" 914
"Ovavisculida" 958
"Metchnikovellida" 541
"Nucleophagales" 982
"Paramicrosporidiales" 1035?
"Morellosporales" 963?
Astathelohaniida 907 # to be edited

Include orders Astathelohaniida (Frizzera et al., 2026), Nosematida, and Areosporida in tree manually and be sure to include in annotations

# Family
"Mrazekiidae" 796
"Enterocytozoonidae" 769
"Encephalitozoonidae" 835
"Nosematidae" 852
"Glugeidae" 641
"Spragueidae" 680
"Unikaryonidae" 702
"Astathelohaniidae" 907
"Neoperezidae" 931
"Tubulinosematidae" 951
"Caudosporidae" 914
"Amblyosporidae" 574
"Gurleyidae" 557
"Amphiacanthidae" 543
"Metchnikovellidae" 541
"outgroup" 1049
"Tuzetidae" 660
"Paramicrosporidiaceae" 1041
"nucleophagaceae" 982
Morellosporaceae 970
Pleistophoridae 3 672
Pleistophoridae 2 661
Pleistophoridae 1 632
Perezidae 621


# Circular tree layout without tip labels 
```{r}
# some accessible pallets 
wacolors$rainier
```
See this link for color-blind friendly colors https://cran.r-project.org/web/packages/khroma/vignettes/tol.html 
```{r}

#order nodes from nexus tree
order <- data.frame(
  node = c(750, 852, 619, 548, 919, 914, 958, 541, 982, 1035, 963, 907),
  fill = c("#44AA99", "#CC6677", "#228833", "#1170AA", "#EE8866", "#88CCEE", "#EF67EB", "#00BD5C", "#882255", "#DDDDDD", "#DDCC77", "#999933"),
  type = c("Enterocytozoonida", "Nosema", "Glugeida", "Amblyosporida", "Neopereziida", "Caudosporida", "Ovavisculida", "Metchnikovellida", "Nucleophagales", "Paramicrosporidiales", "Morellosporales", "Astathelohaniida")
) # edit nosema in .svg file to include encephalitozoonidae

order$type <- factor(order$type, levels = order$type)

order_colors <- setNames(order$fill, order$type)
```

tree with all tip labels
```{r}
full.mic.tree <- tree1 %<+% edit_lb +
  geom_tiplab(size = 2, aes(label = new_lb)) +
  geom_tree(size = .25) +
  geom_treescale() +
  geom_hilight(data = order, 
               mapping = aes(node = node, fill = type),  # Map fill to type
               show.legend = FALSE,
               extendto = 3) +
  scale_fill_manual(values = c("#44AA99", "#CC6677", "#228833", "#1170AA", "#EE8866", "#88CCEE", "#EF67EB", "#00BD5C", "#882255", "#DDDDDD", "#DDCC77", "#999933")) +
  geom_cladelab(data = data.frame( # family
    node = c(796, 769, 835, 852, 641, 632, 661, 672, 680, 702, 621, 907, 931, 951, 914, 574, 557, 543, 541, 1049),
    name = c("Mrazekiidae", "Enterocytozoonidae", "Encephalitozoonidae", "Nosematidae", "Glugeidae", "Pleistophoridae 1", "Pleistophoridae 2", "Pleistophoridae 3", "Spragueidae", "Unikaryonidae", "Perezidae", "Astathelohaniidae", "Neoperezidae", "Tubulinosematidae", "Caudosporidae", "Amblyosporidae", "Gurleyidae", "Amphiacanthidae", "Metchnikovellidae", "outgroup")),
    mapping = aes(node = node,
                  label = name),
    align = TRUE,
    offset = .5,
    barsize = .2,
    fontsize = 2,
    horizontal = TRUE) + 
  new_scale_fill() +
  geom_point2(aes(subset = !isTip & bootstrap == 100),
              shape = 21,
              fill = "black",
              size = 1.5,
              stroke = 0.2) +
  geom_point2(aes(subset = !isTip & bootstrap >= 80 & bootstrap < 100),
              shape = 21,
              fill = "grey",
              size = 1.5,
              stroke = 0.2) 

  ## Theme
  theme_tree()
  
full.mic.tree
```


```{r}
mic.tree_bslabs <- tree2 %<+% edit_lb +
  geom_tiplab(size = 2, aes(label = new_lb)) +
  geom_tree(size = .25) +
  geom_treescale() +
  geom_hilight(data = order, 
               mapping = aes(node = node, fill = type),  # Map fill to type
               show.legend = FALSE,
               extendto = 3) +
  geom_cladelab(data = data.frame(
    node = c(1085, 1059, 1037, 608, 712, 672, 635, 614, 774, 766, 790, 952, 961, 803, 805, 901, 715),
    name = c("Mrazekiidae", "Enterocytozoonidae", "Encephalitozoonidae", "Nosematidae", "Glugeidae", "Spragueidae", "Unikaryonidae", "Astathelohaniidae", "Neoperezidae", "Tubulinosematidae", "Caudosporidae", "Amblyosporidae", "Gurleyidae", "Amphiacanthidae", "Metchnikovellidae", "outgroup", "Tuzetidae")),
    mapping = aes(node = node,
                  label = name),
    align = TRUE,
    offset = .5,
    barsize = .2,
    fontsize = 2,
    horizontal = TRUE) + 
  new_scale_fill() +
#  scale_fill_manual(values = order_colors,  # Use the named vector of colors
#                    guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Orders"),
#                    breaks = order$type) + # Use the type values for breaks 
  geom_text2(aes(subset = !isTip, label = bootstrap),
             size = 2, 
             hjust = 0.5, 
             vjust = -0.5, 
             color = "black")
  theme_tree()
  
mic.tree_bslabs
```

```{r}
ggsave(
  file = "full_mictree.svg",
  plot = full.mic.tree,
  width = 15,
  height = 35)
```
edit .svg file. Subset the tree manually (figure 1)

```{r}

c.tree_simp <- ggtree(raxml, layout = "circular")
c.tree_simp

order.tree <- c.tree_simp +
  geom_hilight(data = order, 
               mapping = aes(node = node, fill = type),  # Map fill to type
               show.legend = FALSE,
               extendto = 4) +
  scale_fill_manual(values = order_colors,  # Use the named vector of colors
                    guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Orders"),
                    breaks = order$type)  # Use the type values for breaks

order.tree
```

Order and family tree with bootstraps
```{r}
clade.tree <- order.tree +
  geom_cladelab(data = data.frame(
        node = c(866, 904, 934, 789, 689, 709, 675, 668, 714, 735, 655, 781, 553, 573, 580, 606, 588, 956, 958, 1041),
    name = c("Mrazekiidae", "Enterocytozoonidae", "Encephalitozoonidae", "Nosematidae", "Glugeidae", "Pleistophoridae 1", "Pleistophoridae 2", "Pleistophoridae 3", "Spragueidae", "Unikaryonidae", "Perezidae", "Astathelohaniidae", "Neoperezidae", "Tubulinosematidae", "Caudosporidae", "Amblyosporidae", "Gurleyidae", "Amphiacanthidae", "Metchnikovellidae", "outgroup")),
    mapping = aes(node = node,
                  label = name),
    align = TRUE,
    offset = 0,
    barsize = .2,
    fontsize = 2,
    horizontal = FALSE) +
  geom_point2(aes(subset = !isTip & bootstrap == 100),
              shape = 21,
              fill = "black",
              size = 1.5,
              stroke = 0.2) +
  geom_point2(aes(subset = !isTip & bootstrap >= 80 & bootstrap < 100),
              shape = 21,
              fill = "grey",
              size = 1.5,
              stroke = 0.2) +
  theme_tree()


clade.tree

```
```{r}
ggsave("clade_tree.svg", 
       clade.tree,
       path = "slides")
```
save this tree to copy and paste the missing order

```{r}
ggsave("order_meta_tree.svg", 
       order.tree,
       path = "slides")
```

# Add metadata
```{r}
metaData <- read.csv("finished_propagated_taxonomy_metadata_moved.csv")

metaData[metaData == ""] <- NA

#Subset metadata to match the tip labels in the tree
metaData_sub <- metaData[metaData$pr2_accession %in% raxml@phylo[["tip.label"]], ]

metaData_sub <- metaData_sub[match(raxml@phylo[["tip.label"]], metaData_sub$pr2_accession), ]

# Addition of metadata by species from Bojko et al. 2022 where previously absent 
b22 <- read.table("C:/Users/south/OneDrive - UBC/Sequences/pr2_workshop/microsporidia/additional_taxa/1000bp_tree/plus_ch_taxa/bojko2022_metadata_select.txt", header = TRUE, sep = "\t")

#for row i in 1:i if environment_type is NA, replace the environment type with that of the corresponding species from b22
for(i in 1:nrow(metaData_sub)) {
  if(is.na(metaData_sub$environment_type[i])) {
    species_match <- metaData_sub$species[i]
    b22_row <- b22[b22$Species == species_match, ]
    if (nrow(b22_row) > 0 ) {
      metaData_sub$environment_type[i] <- b22_row$Environment[1]
    }
  }
}

# do the same for host
for(i in 1:nrow(metaData_sub)) {
  if(is.na(metaData_sub$host_category[i])) {
    species_match <- metaData_sub$species[i]
    b22_row <- b22[b22$Species == species_match, ]
    if (nrow(b22_row) > 0 ) {
      metaData_sub$host_category[i] <- b22_row$Host_group[1]
    }
  }
}

rownames(metaData_sub) <- raxml@phylo[["tip.label"]]

metaData_select <- metaData_sub %>%
  dplyr::select(pr2_sample_type, environment_type, host_category, annotation_note)

metaData_select <- metaData_select %>% 
  mutate(environment_type = tolower(environment_type))

#B22 misspelled "ciliate", so I will replace all "Cilliates" with "Ciliates"
metaData_select$host_category <- gsub("Cilliates", "Ciliates", metaData_select$host_category)

```


```{r}
# Assign Heatmap colors using color blind friendly pallets https://zenodo.org/records/3381072 and wacolors

# Meta Data Tree
meta.tree1 <- order.tree + new_scale_fill()

meta.tree1b <- gheatmap(p = meta.tree1,
           data = metaData_select[,"pr2_sample_type", drop = F],  # Add the heatmap
           color = NULL, 
           colnames_position = "top", 
           colnames_angle = 90, 
           colnames_offset_y = 0.1, 
           hjust = 0, 
           font.size = 2,
           offset = 0,
           width = 0.1) +
  scale_fill_manual(values = c("isolate" = "#E2CD70", #uw$flag
                               "culture" = "#5DB3DB",
                               "environmental" = "#247F5B"), 
                    guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = 'Sequence Source'),
                    breaks = c("isolate", "culture", "environmental"),
                    na.value = "white")  # Use the names of heatmap.colours for breaks


meta.tree2 <- meta.tree1b + new_scale_fill()

meta.tree2b <- gheatmap(meta.tree2, 
                            data = metaData_select[,"environment_type", drop = F],
                            color = NULL, 
                            colnames_position = "top", 
                            colnames_angle = 90, 
                            colnames_offset_y = 1,
                            hjust = 0, 
                            font.size = 2,
                            offset = 0.4,
                            width = 0.1) +
  scale_fill_manual(values = c("brackish" = "#8A6172",
                     "freshwater" = "#69A2E4",
                     "marine" = "#2D3F4A",
                     "terrestrial" = "#94AA3D",
                     "wastewater" = "#8F6D3F"),
                guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Habitat"),
                breaks = c("brackish", "freshwater", "marine", "terrestrial", "wastewater"),
                na.value = "white")

meta.tree3 <- meta.tree2b + new_scale_fill()

meta.tree3b <- gheatmap(meta.tree3, 
                            data = metaData_select[,"host_category", drop = F],
                            color = NULL, 
                            colnames_position = "top", 
                            colnames_angle = 90, 
                            colnames_offset_y = 1,
                            hjust = 0, 
                            font.size = 2,
                            offset = 0.8,
                            width = 0.1) +
  scale_fill_manual(values = c("Annelid" = "#FFAABB",
                               "Arachnids" = "#000000",
                               "Bryozoa" = "#88CCEE",
                               "Cnidaria" = "#CC6677",
                               "Crustacean" = "#882255",
                               "Insects" = "#332288",
                               "Nematodes" = "#EFC519",
                               "Fish" = "#0072B2",
                               "Mammals" = "#D55E00",
                               "Reptiles" = "#117733",
                               "Apicomplexa" = "#F0E442",
                               "Amoebozoa" = "#999933",
                               "Ciliates" = "#DDDDDD",
                               "Planarian" = "#997700"),
                guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Host"),
                breaks = c("Arachnids", "Crustacean", "Nematodes", "Apicomplexa", "Ciliates", "Insects", "Amoebozoa", "Mammals", "Reptiles", "Bryozoa", "Fish", "Cnidaria", "Annelid", "Planarian"),
                na.value = "white")


meta.tree4 <- meta.tree3b + new_scale_fill()

meta.tree4b<- gheatmap(meta.tree4, 
                            data = metaData_select[,"annotation_note", drop = F],
                            color = NULL, 
                            colnames_position = "top", 
                            colnames_angle = 90, 
                            colnames_offset_y = 1,
                            hjust = 0, 
                            font.size = 2,
                            offset = 1.2,
                            width = 0.1) +
  scale_fill_manual(values = c("corrected" = "#CC7810",
                     "improved" = "#EFC519",
                     "introduced" = "#831285"), #wacolor$skagit
                guide = guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Annotation note"),
                breaks = c("corrected", "improved", "introduced"),
                na.value = "white")

meta.tree4b
```
#save the metadata from this tree to easily add to existing .svg
```{r}
ggsave("no_family_meta_tree.svg", 
       meta.tree4b,
       path = "path")
```



# Add family level labels, branch support, and scale bar to tree


```{r}
family.tree <- meta.tree4b +
  geom_cladelab(data = data.frame(
        node = c(796, 769, 835, 852, 641, 632, 661, 672, 680, 702, 621, 907, 931, 951, 914, 574, 557, 543, 541, 1049),
    name = c("Mrazekiidae", "Enterocytozoonidae", "Encephalitozoonidae", "Nosematidae", "Glugeidae", "Pleistophoridae 1", "Pleistophoridae 2", "Pleistophoridae 3", "Spragueidae", "Unikaryonidae", "Perezidae", "Astathelohaniidae", "Neoperezidae", "Tubulinosematidae", "Caudosporidae", "Amblyosporidae", "Gurleyidae", "Amphiacanthidae", "Metchnikovellidae", "outgroup")),
    mapping = aes(node = node,
                  label = name),
    align = TRUE,
    offset = 1.8,
    barsize = .2,
    fontsize = 2,
    horizontal = FALSE) +
  theme_tree()


family.tree
```

```{r}
# save file
ggsave("meta_tree.svg", 
       family.tree)

ggsave("tree.png", 
       family.tree,
       path = "path")
```

# Generate a simpler Tree for talks
```{r}
order.tree.simp <- meta.tree4 +
  geom_cladelab(data = order,
    mapping = aes(node = node,
                  label = type),
    align = TRUE,
    offset = 1,
    barsize = .2,
    fontsize = 2,
    horizontal = FALSE) +
  new_scale_fill() +
  geom_point2(aes(subset = !isTip,
                  fill = cut(bootstrap, c(0, 70, 90, 100))),
              shape = 21,
              size = .5,
              stroke = 0.01) +
  theme_tree() +
  scale_fill_manual(values=c("white", "darkgrey", "black"),
                    guide=guide_legend(keywidth = 0.5, keyheight = 0.5, title = "Bootstrap Support"),  
                    breaks=c('(0,70]', '(70,90]', '(90,100]'), 
                    labels=expression(BP < 70, 70 <= BP * " < 90", BP>=90)) +
  geom_treescale(fontsize = 2, offset = 1)

order.tree.simp
```

```{r}
# save file
ggsave("simp_tree.svg", 
       order.tree.simp,
       path = "path")

ggsave("simp_tree.png", 
       order.tree.simp,
       path = "path")
```






